Fixing Vulnerabilities Hinders Maintainability

### 2019-01-31 paper/ICPC19/icpc19.tex
small advances

Software quality is important because it is ultimately related to the overall
cost for developing, extending and maintaing software applications. Software
quality characteristics include, but are not limited to, functional correctness,
reliability, usability, maintainability, and security. Security is an important
non-functional requirement during the development of software systems.  However,
there is still a considerable amount of zero-day vulnerabilities being
discovered, and fixed, almost weekly as disclosed by the Zero Day Initiative
website\footnote{Zero Day Initiative website available  at
https://www.zerodayinitiative.com/advisories/published/ (Accessed on \today{})}.
As an example, in the last week of January 2019, a severe vulnerability in Apple
FaceTime was
revealed\footnote{https://thehackernews.com/2019/01/apple-facetime-privacy-hack.html
(Accessed on \today{})}. These vulnerabilities are gateways for attackers to
penetrate systems and exploit their resources.

% Maintainability importance is being understimated.

% The time required to implement an additional security functionality in an application is not always easy to estimate. Estimating whether such functionality requires any changes in the architecture of the system is even harder. For instance, if you think about refining the access control policies of your application, by moving from a function-level access control to a data-level access control (where access control is performed on the data field instead on the system functionality executed by the user), can you precisely estimate the amount of time required to perform such change or even if that change has an impact on the architecture of the system?

% Having a resilient and scalable security architecture and design is crucial to prevent vulnerabilities.

% High complex code is hard to test, thus being more difficult to find bugs and vulnerabilities. Therefore, writing maitainable code prevent the introduction of vulnerabilities in the code.

@@

@@

>>>>

Software quality is important because it is ultimately related to the overall
cost for developing, extending and maintaing software applications. Software
quality characteristics include, but are not limited to, functional correctness,
reliability, usability, maintainability, and security. Security is an important
non-functional requirement during the development of software systems.  However,
there is still a considerable amount of zero-day vulnerabilities being
discovered, and fixed, almost weekly as disclosed by the Zero Day Initiative
website\footnote{Zero Day Initiative website available  at
https://www.zerodayinitiative.com/advisories/published/ (Accessed on \today{})}.
As an example, in the last week of January 2019, a severe vulnerability in Apple
FaceTime was
revealed\footnote{https://thehackernews.com/2019/01/apple-facetime-privacy-hack.html
(Accessed on \today{})}. Callers could see and hear others without them picking up the call. These vulnerabilities are gateways for attackers to
penetrate systems and exploit their resources.

@@

% find work about the struggle to make systems secure

We believe developers should be able to create secure code without affecting 
the maintainability of their projects.

\begin{itemize}
	\item \textbf{Security Frameworks.} The most popular languagues (e.g., Java and c\#) 
used in the development of enterprise applications provide stable
frameworks to implement the security requirements and easily prevent some of the
OWASP TOP 10 vulnerabilities.
	\item \textbf{Documentation}
	\item \textbf{Programming Languages}
	\item \textbf{Proper authorization mechanisms}
	\item \textbf{Remove dead and duplicated code}
	\item \textbf{Configurations in files}
	\item \textbf{Framework manages the session}

% #1 Security frameworks are your new best friend
% The most popular software stacks (e.g.: Java and C#) used in the development of enterprise applications provide very good and stable frameworks to implement the security requirements and easily prevent some of the OWASP TOP 10 vulnerabilities. For instance, through a very simple annotation such as @EnableWebSecurity, Spring Security automatically and out of the box, implements mitigations against some of the most common web application vulnerabilities (e.g.: CSRF, XSS, etc.). The essential HTTP headers to prevent certain types of vulnerabilities are added to the requests and validated in the responses (e.g.: same-origin policy for the synchronization token to prevent CSRF).
%
% Although these frameworks may require some time to understand and use (the learning curve might be quite steep, sometimes), they have been improving quite a lot in the past few years, facilitating the implementation and maintenance of the security requirements.
%
% Nevertheless, do not blindly trust all the defaults of the frameworks and libraries your application depends on. More than often frameworks rely on weak default encryption algorithms such as DES (encryption standard deprecated since 1998). Thus, always check all the frameworks and libraries defaults, before releasing the application.
%
% Besides, some of the least maintained frameworks contain vulnerabilities that never get to be reported. This is because some bugs (this is an interesting discussion about a systemd null pointer dereferencing bug) are not even considered security vulnerabilities in the first place. Notice that categorisation of bugs as security vulnerabilities, is based on the ability of the development team to evaluate the impact of that bug in the security of the application. That’s why updating the libraries and frameworks of your application to their latest version helps improving its overall security. So, stay up-to-date.
%
% #2 Deny first, authorize after
% Broken access control is one of the OWASP TOP 10 vulnerabilities. This is because most applications fail to implement proper authorization mechanisms. Authorization design and implementation is probably one of the most complex security features that exists. Making it right and scalable is challenging and requires the effort of all the team members. Nevertheless, adoption of certain mechanisms and design strategies may facilitate authorization maintenance in the long run:
%
% 1. Authorise only afterwards: access must always be denied in the first place, i.e., access must only be granted if the user is actually authorized to perform a certain action. So, access control policies must be checked before processing the request and not when producing the response.
%
% 2. Use and abuse of filters: HTTP filters can be used to filter out all the requests before processing them (e.g.: if the user doing the request is authorized to perform the designated action, then the request can be processed). Certain frameworks (e.g.: Spring Security) even provide a very elegant way of binding the access control policies to certain system functions by means of code annotations in the methods that expose the desired functionalities. In combination with the HTTP filters these annotations provide a very clean and easy to change implementation of the access control mechanisms.
%
% 3. Always document what you’re doing: either using role-based access control or a more fine-grained access control model, the functionalities enabled in each roles must be clearly defined and documented. Although, the code should speak for itself, the lack of documentation brings risks for code portability and transferability, specially when considering access control management.
%
% #3 Make the IT infrastructure your number one software project
% A securely developed application is defenseless against an attacker with administrative privileges in the production environment. For instance, the WannaCry ransomware attack was based on an OS level exploit that was fairly easy to protect, just by upgrading the OS version. Nonetheless, it was very problematic because a lot of IT departments tend to disregard the importance and impact of OS security updates.
%
% In general, the longer you wait to upgrade, the harder it gets and eventually a system can get ‘un-upgradable’, therefore, unmaintainable.
% When deploying the application, it is necessary to make sure that all 3rd party components that are part of the production environment can be upgraded or recreated from backup (in case of ransomware) at any point in time. This includes databases and web servers, deployment frameworks, the identity manager, log manager application and alerting system, the operating system, etc.
%
% The ‘upgrading is hard’ problem is similar to the ‘releasing is hard’ problem and should be solved the same way: by doing it often.
% The way to do this is to treat the environment in which your application runs as part of the application and not something separate. As such it should be treated as a software project in itself.
%
% #4 Validations should be done in the backend
% More and more web applications follow a service oriented architecture model, where the backend provides a RESTful API and the front-end is purely implemented in JavaScript (with HTML). Some modern JavaScript frameworks provide out of the box escaping and sanitization functionalities (e.g.: Angular 2 or above). Some developers are led to mistakenly believe that this free-feature, together with type checking the REST parameters (done in the corresponding controller, for instance) is enough to prevent introduction of malicious input in the application. But unfortunately, it is not. As much as the web and security frameworks try to enable security mitigations by default, a lot of effort is still needed from the development team. Therefore, it’s very important to validate all the user input in the backend and filter-out unwanted/invalid inputs.
%management of the web application.
%
% Notice that the fact that in modern frameworks crypto primitives can be defined in a configuration file, makes it much easier and less time consuming, to, for instance, update those primitives when a new attack is discovered. Plus, if the default session manager has a known vulnerability, it is easier to patch the application by updating the library, then doing quick (and perhaps dirty) fixes in custom implementations, that can lead to even more exploits.
%

@@

\subsection{Security Maintainability Models}

~\cite{6616351,oswap:2009,common:2009}
